Hunting experience shapes individual foraging specialisation and predator-prey interactions in an online videogame

### 2023-02-21 manuscript/Chapter2-manuscript.Rmd
include prior description mdhglm

where $env$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their incidence matrices $M$, $N$, and $Z$. $b$ is the vector of fixed effects with its incidence matrix $X$. The $m$ and $d$ subscripts indicate that an estimate is from the mean or dispersion part of the model, respectively. We assumed that all random effects followed a Gaussian distribution with estimated standard deviation ($env \sim N(0,\ I\ \sigma_{env})$, $av \sim N(0,\ I\ \sigma_{av})$, and $id \sim N(0,\ I\ \sigma_{id})$). The parameters $\sigma_{env}$, $\sigma_{av}$, and $\sigma_{id}$, correspond to the among environment, among avatar, and among individual standard deviations, respectively. $I$ is the identity matrix, with $I\sigma$ indicating that the random effects are independently and identically distributed. Here, we assume that the residuals vary among individuals where $e \sim N(0,\ Diag\ [\sigma_{e}])$, $[\begin{array}{cc} id_m \\ id_v \end{array}] \sim N\left(\begin{array}{cc} 0 \\ \\ 0 \end{array},\ \Sigma_{id} \otimes\ I \right)$, and $\Sigma_{id}$ is a variance covariance matrix of the individual identity random effect. Equations (1) and (2) were parametrized with a Gaussian distribution family (predator and prey speed), while equation (3) was parametrized with a beta-binomial distribution family (predator hunting success). We parametrized the model to run four MCMC chains with 1000 posterior samples for each parameter. To do so, we ran 2500 iterations with a thinning set to 8, with the first 500 iterations used as warmups. We explain how we specified the priors of the MDHGLM in the supporting information.

We estimated the correlation between the individual random effect on the mean and on the dispersion within each trait (e.g. $cor(\sigma_{id_{m}\ pred}\ \sigma_{id_{d}\ pred})$) and among each trait (e.g. $cor(\sigma_{id_{m}\ pred}\ \sigma_{id_{d}\ prey})$) within a level of experience. In the first case, this allows us to assess, for example, whether predators that are on average slower/faster are more specialized or flexible. In the second case, the correlations enables us to estimate, for example, whether predators that encountered prey that were faster/slower adopted a slower/faster (or more specialized/flexible) tactic, or if slower/faster predators had lower/greater hunting success. We also evaluated these correlations among the different levels of experience (e.g. $cor(\sigma_{id_{m}\ pred\ novice}\ \sigma_{id_{d}\ pred\ intermediate})$). The latter "character-state" approach [@Via.Lande1985;@viaAdaptivePhenotypicPlasticity1995], enables us to investigate whether predators express similar behaviours across each developmental stage. The correlations were estimated assuming a multivariate Gaussian distribution. We also report additional analyses, such as variance partitioning, and coefficients of variation, in the supporting information.

@@

For decades, hunting experience has been hypothesized to play a major role in predator-prey interactions []. Yet, the lack of empirical evidence has led the existing literature to predict conflicting outcomes []. It has thus remained unclear whether predators specialize in their tactic with experience, how such specialization emerges depending on the type of prey that predators encounter, and whether specialization with experience leads to an increase in hunting success. We adressed these questions by capitalizing on an online videogame system simulating a predator-prey interaction, where both the predator and the prey behaviours are simultaneously monitored. Our results show that experience does not generate responses at the population level. Instead, experience appears to be a property of individuals, such that each predator developed, through extensive practice, their own behavioural trajectory. Individuals became increasingly distinct hunters with experience; some specialized at fast-paced hunting, while others adopted a slower and more flexible tactic. The predators' tactics were also associated with prey encounters, such that those that adopted a specialized cursorial tactic experienced more predictable encounters, while slower and flexible hunters experienced more unpredictable encounters, with no difference in success between the two types of hunters. With experience, this predator-prey behavioural relationship grew stronger, while the relationship between the prey's behaviour and hunting success diminished. Together, our observations suggest that predators use their experience to adjust their hunting tactic, and the degree at which they specialize in it, based on the prey that they encountered throughout their lifetime. Moreover, the stronger predator-prey interaction through experience may imply the emergence of an arms race mediated through learning.

>>>>

where $env$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their incidence matrices $M$, $N$, and $Z$. $b$ is the vector of fixed effects with its incidence matrix $X$. The $m$ and $d$ subscripts indicate that an estimate is from the mean or dispersion part of the model, respectively. We assumed that all random effects followed a Gaussian distribution with estimated standard deviation ($env \sim N(0,\ I\ \sigma_{env})$, $av \sim N(0,\ I\ \sigma_{av})$, and $id \sim N(0,\ I\ \sigma_{id})$). The parameters $\sigma_{env}$, $\sigma_{av}$, and $\sigma_{id}$, correspond to the among environment, among avatar, and among individual standard deviations, respectively. $I$ is the identity matrix, with $I\sigma$ indicating that the random effects are independently and identically distributed. Here, we assume that the residuals vary among individuals where $e \sim N(0,\ Diag\ [\sigma_{e}])$, $[\begin{array}{cc} id_m \\ id_v \end{array}] \sim N\left(\begin{array}{cc} 0 \\ \\ 0 \end{array},\ \Sigma_{id} \otimes\ I \right)$, and $\Sigma_{id}$ is a variance covariance matrix of the individual identity random effect.

We parameterized equations (1) and (2) with a Gaussian distribution family (predator and prey speed), and employed a beta-binomial distribution family for equation (3) (predator hunting success). We employed Gaussian priors for the prey rank ($N(2.5,\ 1)$) and game duration ($N(10,\ 2)$) on the mean part of the model, and the default priors in "brms" for the dispersion part. We used the default Student t priors in "brms" for the intercepts on the mean and dispersion part of the equation. For hunting success, we applied a Gaussian prior on the precision parameter ($N(2,\ 1)$). For every predicted variable, we applied a half-Gaussian on the random effects for the mean part of the model ($N(0,\ 1)$), and the default "brms" Student t prior for the dispersion part of the model. We parametrized the model to run four MCMC chains with 1000 posterior samples for each parameter. To do so, we ran 2500 iterations with a thinning set to 8, with the first 500 iterations used as warmups.

We estimated the correlation between the individual random effect on the mean and on the dispersion within each trait (e.g. $cor(\sigma_{id_{m}\ pred}\ \sigma_{id_{d}\ pred})$) and among each trait (e.g. $cor(\sigma_{id_{m}\ pred}\ \sigma_{id_{d}\ prey})$) within a level of experience. In the first case, this allows us to assess, for example, whether predators that are on average slower/faster are more specialized or flexible. In the second case, the correlations enables us to estimate, for example, whether predators that encountered prey that were faster/slower adopted a slower/faster (or more specialized/flexible) tactic, or if slower/faster predators had lower/greater hunting success. We also evaluated these correlations among the different levels of experience (e.g. $cor(\sigma_{id_{m}\ pred\ novice}\ \sigma_{id_{d}\ pred\ intermediate})$). The latter "character-state" approach [@Via.Lande1985;@viaAdaptivePhenotypicPlasticity1995], enables us to investigate whether predators express similar behaviours across each developmental stage. The correlations were estimated assuming a multivariate Gaussian distribution. We additionaly report variance partitioning and coefficients of variations in the supporting information.

@@

Hunting experience has long been hypothesized to play an important role in predator-prey interactions []. Yet, the lack of empirical evidence has led the existing literature to predict conflicting outcomes []. It has thus remained unclear whether predators specialize in their tactic with experience, how such specialization emerges depending on the type of prey that predators encounter, and whether specialization with experience leads to an increase in hunting success. We adressed these questions by capitalizing on an online videogame system simulating a predator-prey interaction, where both the predator and the prey behaviours are simultaneously monitored. Our results show that experience does not generate responses at the population level. Instead, experience appears to be a property of individuals, such that each predator developed, through extensive practice, their own behavioural trajectory. Individuals became increasingly distinct hunters with experience; some specialized at fast-paced hunting, while others adopted a slower and more flexible tactic. The predators' tactics were also associated with prey encounters, such that those that adopted a specialized cursorial tactic experienced more predictable encounters, while slower and flexible hunters experienced more unpredictable encounters, with no difference in success between the two types of hunters. With experience, this predator-prey behavioural relationship grew stronger, while the relationship between the prey's behaviour and hunting success diminished. Together, our observations suggest that predators use their experience to adjust their hunting tactic, and the degree at which they specialize in it, based on the prey that they encountered throughout their lifetime. Moreover, the stronger predator-prey interaction through experience may imply the emergence of an arms race mediated through learning.
